---
title: "Game Test"
author: "Rob Carnell and Nancy McMillan"
date: "April 13, 2021"
output: html_document
---

```{r setup, include=FALSE}
set.seed(193932)
knitr::opts_chunk$set(echo = TRUE)

require(reticulate)
reticulate::use_python("C:/ProgramData/Anaconda3/python.exe")
require(numDeriv)
require(mvtnorm)
```

## Notation

- $X$ multivariate random variable representing the true distribution.
- $P$ is the set of parameters for the true distribution.  
- $Y$ multivariate random variable for the adversary distribution
- $n_A$ the number of adversaries
- $\theta$ set of parameters for the adversary distribution
- $z$ multivariate test value or bracket
- $s(z|X=x)$ is the score of the test bracket against a bracket from the truth 
distribution
- $s(Y_k|X=x)$ is a random variable of scores with a distribution induced by the 
adversary distribution given a bracket from the truth distribution, where $k=1,\ldots,n_A$ adversaries
- $f$ is the distribution function of $X$

The goal is to maximize: $\int P(s(z|X=x) > s(Y_k|X=x) \forall k) f(x) dx$

## Centered True and Crowd distributions with 5 adversaries

```{r}
set.seed(184490)

Nsims <- 1000
Nplayers <- 5
meanx <- c(1,1)
Sigmax <- matrix(c(2,0.2,0.2,2), nrow = 2)
meany <- c(1,1)
Sigmay <- matrix(c(1,-0.1,-0.1,1), nrow = 2)
# truth
X <- rmvnorm(Nsims, meanx, Sigmax)
#apply(X, 2, mean)

# crowd
Y <- array(rmvnorm(Nsims*Nplayers, meany, Sigmay), dim = c(Nsims, length(meany), Nplayers))
#apply(Y, c(2,3), mean)

s <- function(u,v) sum((u - v)^2)
#s(c(2,3),c(4,5))

require(ggplot2)
ggplot(data.frame(x1 = X[,1], x2 = X[,2]), aes(x = x1, y = x2)) + 
  geom_density_2d(col = "blue") +
  geom_density_2d(aes(x = x1, y = x2), data = data.frame(x1 = Y[,1,1], x2 = Y[,2,1]), col = "red") +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 1, z2 = 1), col = "red", size = 3) +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 1, z2 = 1), col = "blue", size = 1)

```

### Probability of Winning

```{r}
f <- function(z)
{
  #z <- c(0,0)
  sz <- apply(X, 1, function(x) s(x,z))
  sy <- sapply(1:Nsims, function(n) {sapply(1:Nplayers, function(m) {s(X[n,], Y[n,,m])})})
  ind <- which(sapply(1:Nsims, function(i) all(sz[i] < sy[,i])))
  # % of the time winning
  return(length(ind) / Nsims)
}
zgrid <- data.frame(expand.grid(z1 = seq(-5, 5, by = 0.5), z2 = seq(-5, 5, by = 0.5)))
zgrid$p <- apply(zgrid, 1, f)

#zgrid[which.max(zgrid$p),]

require(ggplot2)
ggplot(zgrid, aes(x = z1, y = z2, z = p, col = p)) + 
  geom_point(size = 2) + 
  geom_contour() +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 1, z2 = 1, p = 1), col = "red", size = 3) +
  geom_point(aes(x = z1, y = z2), data = zgrid[which.max(zgrid$p),], col = "green", size = 3)

require(magrittr)
require(dplyr)
ggplot(zgrid %>% dplyr::filter(z1 == 1), aes(x = z2, y = p)) + 
  geom_line() +
  geom_vline(xintercept = 1, col = "red") +
  geom_hline(yintercept = max(zgrid$p), col = "green")
ggplot(zgrid %>% dplyr::filter(z2 == 1), aes(x = z1, y = p)) + 
  geom_line() +
  geom_vline(xintercept = 1, col = "red") +
  geom_hline(yintercept = max(zgrid$p), col = "green")
```


## Non-centered True and Crowd distributions with 5 adversaries

```{r}
set.seed(3843484)

Nsims <- 1000
Nplayers <- 5
meanx <- c(0,0)
Sigmax <- matrix(c(2,0.2,0.2,2), nrow = 2)
meany <- c(1,1)
Sigmay <- matrix(c(1,-0.1,-0.1,1), nrow = 2)
X <- rmvnorm(Nsims, meanx, Sigmax)
Y <- array(rmvnorm(Nsims*Nplayers, meany, Sigmay), dim = c(Nsims, length(meany), Nplayers))
ggplot(data.frame(x1 = X[,1], x2 = X[,2]), aes(x = x1, y = x2)) + 
  geom_density_2d(col = "blue") +
  geom_density_2d(aes(x = x1, y = x2), data = data.frame(x1 = Y[,1,1], x2 = Y[,2,1]), col = "red") +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 1, z2 = 1), col = "red", size = 3) +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 0, z2 = 0), col = "blue", size = 3)
```

### Probability of Winning

```{r}
zgrid <- data.frame(expand.grid(z1 = seq(-5, 5, by = 0.5), z2 = seq(-5, 5, by = 0.5)))
zgrid$p <- apply(zgrid, 1, f)

ggplot(zgrid, aes(x = z1, y = z2, z = p, col = p)) + 
  geom_point(size = 2) + 
  geom_contour() +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 1, z2 = 1, p = 1), col = "red", size = 3) +
  geom_point(aes(x = z1, y = z2), data = data.frame(z1 = 0, z2 = 0, p = 1), col = "blue", size = 3) +
  geom_point(aes(x = z1, y = z2), data = zgrid[which.max(zgrid$p),], col = "green", size = 3)

ggplot(zgrid %>% dplyr::filter(z1 == 0), aes(x = z2, y = p)) + 
  geom_line() +
  geom_vline(xintercept = 0, col = "blue") +
  geom_vline(xintercept = 1, col = "red") +
  geom_hline(yintercept = max(zgrid$p), col = "green")
ggplot(zgrid %>% dplyr::filter(z2 == 0), aes(x = z1, y = p)) + 
  geom_line() +
  geom_vline(xintercept = 0, col = "blue") +
  geom_vline(xintercept = 1, col = "red") +
  geom_hline(yintercept = max(zgrid$p), col = "green")
```


