---
title: "Bracket Optimization"
author: "RC"
date: "5/12/2021"
output: html_document
params:
  output_file: default.csv
  start_point: Xmax
  elite8_team: Blank
  n_truth_brackets: 100
  n_adversaries: 5
  pRate: 3
  nchildren: 10
  ngenerations: 5
  gen_method: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(ggplot2)
require(dplyr)

source("../utility_functions.R")
load("../2021BracketMetadata.Rdata")

# simulation parameters
n_truth_brackets <- params$n_truth_brackets
n_adversaries <- params$n_adversaries

# genetic algorithm parameters
pRate <- params$pRate
nchildren <- params$nchildren
ngenerations <- params$ngenerations
gen_method <- params$gen_method

# common simulation data
set.seed(193849)

X <- drawTruthSampleList(P, structureList, n_truth_brackets)
checkTruthSampleList(X, structureList)

Y <- drawAdversarySampleList(Th, structureList, n_truth_brackets, n_adversaries)
checkAdversarySampleList(Y, structureList, n_truth_brackets, n_adversaries)

Yscore <- scoreAdversarySampleList(Y, X, n_truth_brackets)
```

## Report Parameters

- Output File: `r params$output_file`
- Type of Start Point: `r params$start_point`
- Elite 8 Team: `r params$elite8_team`
- Number of truth Brackets: `r params$n_truth_brakcets`
- Number of Adversaries: `r params$n_adversaries`
- Poisson Mutation rate: `r params$pRate`
- Number of children per generation: `r params$nchildren`
- Number of generations: `r params$ngenerations`
- Bracket permutation method: `r params$gen_method`

## Create the Starting Bracket

```{r createStartBracket}
if (params$start_point == "Xmax")
{
  Z <- drawMaxLikelihoodBracket(structureList, P)
} else if (params$start_point == "Ymax")
{
  Z <- drawMaxLikelihoodBracket(structureList, Th)
} else if (params$start_point == "Elite8")
{
  team_ind <- which(trans$TEAM == params$elite8_team)
  assertthat::assert_that(length(team_ind) > 0,
                          msg = paste0("Team ", params$elite8_team, " not found"))
  Z <- drawMaxLikelihoodBracket(structureList, P, team_ind)
} else
{
  stop(paste0("Algorithm Start Point ", params$start_point, " not recognized"))
}
checkBracket(structureList, Z)
Zscore <- scoreTruthSampleList(Z, X)
```

## Genetic Algorithm

```{r geneticAlgorithm}
newZ <- geneticBracketSearch(Z, pRate, nchildren, ngenerations, TRUE, gen_method)

dimnames(newZ$current_best) <- list(trans$TEAM, paste0("Round", 1:6))
write.csv(newZ$current_best, file = params$output_file)
```

- Score went from `r newZ$start_score` to `r newZ$score`
- P(win) went from `r newZ$start_pwin` to `r newZ$pwin`

## Score Density

```{r score-density, echo=FALSE}
plot_data <- data.frame(
  group = c(rep("Original Bracket", length(Zscore)),
            rep("Adversary Brackets", length(unlist(Yscore))),
            rep("New Bracket", length(newZ$scores))),
  score = c(Zscore, unlist(Yscore), newZ$scores)
)

ggplot(plot_data, aes(x = score, group = group, col = group)) +
  geom_density() +
  labs(x = "Score", y = "Density", col = "") +
  xlim(0, 1920) +
  geom_vline(aes(xintercept = score, col = group), lty = 2, 
             data = plot_data %>% group_by(group) %>% summarise(score = mean(score)))
```

## Bracket Fingerprint

```{r bracket-image, echo=FALSE, fig.height=8}
plot_data2 <- data.frame(
  group = factor(rep(c("Original Bracket", "New Bracket", "Mean Adversary", "Mean Truth"), 
                     each = length(trans$TEAM)),
                 levels = c("Mean Truth", "Mean Adversary", "Original Bracket", "New Bracket")),
  team = factor(rep(trans$TEAM, times = 4), levels = rev(trans$TEAM)),
  round = c(rowSums(Z), 
            rowSums(newZ$current_best), 
            apply(sapply(Y, function(z) {
                              apply(sapply(z, rowSums), 1, mean)
                            }), 1, mean),
            apply(sapply(X, function(z) rowSums(z)), 1, mean))
)

ggplot(plot_data2, aes(x = round, y = team, fill = group, group = group)) +
  geom_bar(stat = "identity", width = 0.5, position = position_dodge()) +
  facet_grid(. ~ group) +
  labs(y = "", x = "Round", fill = "")

```
